### 2024-04-27 thesis/chapters/analysis/2-features.tex
Export graphs from experiments and prepare outline of thesis

It is important to note that the design of features is not a standalone step in the machine learning pipeline, but it should be done iteratively to improve the target model. Signal features are computed in the temporal and spectral domain~\cite{brito_fault_2021}.

\subsection{Temporal domain features}
The most widely found features in the literature are rudimentary statistical measures of the central moment: mean, variance, standard deviation, skewness, and kurtosis~(Tab.~\ref{tab:td-features}). Statistics can be calculated in any domain, but the mean value should not be used in detrended data. The vibration severity metrics out of technical standards are also highly regarded. The amplitude characteristics include root-mean-square and peak-to-peak distance~\cite{mostafavi_novel_2021}. The waveform shape is captured by zero-crossing rate and average amplitude change~\cite{moctar_time-domain_2023}.

@@

\begin{table}[h]
\centering
\renewcommand{\arraystretch}{2}
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Feature}            & \textbf{Equation}                                                                    \\ \hline
\textit{Peak-to-peak}       & $ X_{ppv} = \max(x_i) - \min(x_i) $               
\\ \hline
\textit{Zero-crossing rate} & $X_{zc} = \frac{1}{N}\sum_{i = 1}^{N}{\mathrm{sgn}(x_i \cdot x_{i-1})} $
\\ \hline
\textit{Root mean square}   & $ X_{rms} = \sqrt{\frac{1}{N}\sum_{i = 1}^{N}{x_i^2}} $ 
\\ \hline
\textit{Skewness}           & $ X_{sv} = \frac{1}{N}\sum_{i = 1}^{N}{\left(\frac{x_i - \bar{x}}{\sigma}\right)^3}$      
\\ \hline
\textit{Kurtosis}           & $ X_{kv} = \frac{1}{N}\sum_{i = 1}^{N}{\left(\frac{x_i - \bar{x}}{\sigma}\right)^4}$                                           
\\ \hline
\textit{Shape factor}   & $ X_{sf} = \frac{X_{rms}}{\frac{1}{N} \sum_{i=1}^{N}{|x_i|}}$  
\\ \hline
\textit{Crest factor}   & $ X_{cf} = \frac{max(|x_i|)}{X_{rms}} $                           \\ \hline
\textit{Impulse factor} & $ X_{if} = \frac{max(|x_i|)}{\frac{1}{N} \sum_{i=1}^{N}{|x_i|}} $                                                 
\\ \hline       
\textit{Clearance factor}  & $ X_{mf} = \frac{max(|x_i|)}{\left( \frac{1}{N} \sum_{i=1}^{N}{\sqrt{|x_i|}} \right)^2} $ 
\\ \hline
\textit{Average Amplitude Change (AAC)}        & $ X_{AAC} = \frac{1}{N}\sum_{i = 1}^{N - 1}{|x_{i + 1} - x_i|} $                             \\ \hline                                                                                     
\end{tabular}
\caption{Temporal domain features}
\label{tab:td-features}
\end{table}

\subsection{Spectral domain features}
The mechanical faults present themselves as oscillatory patterns which are combinations of frequencies with various amplitudes. The Fourier transform is one of the most prominent strategies in power spectral density estimation. Experts on vibrodiagnostics utilize it as a primary signal processing technique for data analysis as it is recommended in ISO 13373-2 standard~\cite{noauthor_iso_2016_2}.

@@

\begin{table}[ht]
\renewcommand{\arraystretch}{2}
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Feature}           & \textbf{Equation}                                                                                                  \\ \hline
\textit{Spectral centroid} & $ X_{fc} = \frac{\sum_{i = 0}^{N - 1}{f_i \cdot a(f_i)}}{\sum_{i = 0}^{N - 1}{a(f_i)}}$                   \\ \hline 
\textit{Energy}            & $ E(N) = \sum_{i = 1}^{N} a^2(t) $                                                                    \\ \hline \textit{Energy ratio}                & $E_r = E_i \;/\; \sum_{i = 1}^{N}{E_i} $                                                        \\ \hline
\textit{Spectral roll-on} & $ E(f_c) = 0.05 \cdot E(f_s / 2) $ \\ \hline                    
\textit{Spectral roll-off} & $ E(f_c) = 0.85 \cdot E(f_s / 2) $                                                   \\ \hline
\textit{Spectral flux}     & $X_{\mathrm{flux}} = 1 - \mathrm{corr}(A_{t-1}, A_t)$ \\ \hline 
\textit{Noisiness}                   & $X_{noise} = \mathrm{SNR}(X) = \mu \;/\; \sigma $                             \\ \hline
\textit{Shannon entropy}      & $H(X) = - \sum_{i} P(X = x_i) \cdot \ln(P(X = x_i)) $                                                  \\ \hline
\textit{Spectral negentropy}      & $\Delta I_E(f, \Delta f) = \sum_{k = 0}^{N - 1}{\frac{E(k, f, \Delta f)^2}{\frac{1}{N} \sum_{k=0}^{N-1} E(k, f, \Delta f)^2}} \cdot \ln\left(\frac{E(k, f, \Delta f)^2}{\frac{1}{N} \sum_{k=0}^{N-1} E(k, f, \Delta f)^2}\right)$                                                \\ \hline
\end{tabular}
\caption{Frequency domain features}
\label{tab:fd-features}
\end{table}

In the spectral domain, we can obtain the usual statistical properties of the distribution being spectral centroid, skew and kurtosis. Additionally, spectral roll-on and roll-off, fundamental frequency, entropy, negentropy, spectral flux, signal-to-noise ratio (noisiness), energy in frequency bands, and energy ratio are extracted (Tab.~\ref{tab:fd-features})~\cite{peeters_large_2004}.

@@

@@

\subsection{Scores for ranking}
Most of the scores are based on supervised learning, so they expect true class labels to apportion the measurements respectively. After the scores are assigned to the first $n$ features, those below a threshold are removed.

>>>>

It is important to note that the design of features is not a standalone step in the machine learning pipeline, but it should be done iteratively to improve the target model. Signal features are computed in the time and frequency domain~\cite{brito_fault_2021}.

\subsection{Time-domain features}
The most widely found features in the literature are rudimentary statistical measures of the central moment: mean, variance, standard deviation, skewness, and kurtosis~(Tab.~\ref{tab:td-features}). Statistics can be calculated in any domain, but the mean value should not be used in detrended data. The vibration severity metrics out of technical standards are also highly regarded. The amplitude characteristics include root-mean-square and peak-to-peak distance~\cite{mostafavi_novel_2021}. The waveform shape is captured by zero-crossing rate and average amplitude change~\cite{moctar_time-domain_2023}.

@@

\begin{table}[h]
\centering
\renewcommand{\arraystretch}{2}
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Feature}            & \textbf{Equation}                                                                    \\ \hline
\textit{Peak-to-peak}       & $ X_{ppv} = \max(x_i) - \min(x_i) $               
\\ \hline
\textit{Zero-crossing rate} & $X_{zc} = \frac{1}{N}\sum_{i = 1}^{N}{\mathrm{sgn}(x_i \cdot x_{i-1})} $
\\ \hline
\textit{Root mean square}   & $ X_{rms} = \sqrt{\frac{1}{N}\sum_{i = 1}^{N}{x_i^2}} $ 
\\ \hline
\textit{Skewness}           & $ X_{sv} = \frac{1}{N}\sum_{i = 1}^{N}{\left(\frac{x_i - \bar{x}}{\sigma}\right)^3}$      
\\ \hline
\textit{Kurtosis}           & $ X_{kv} = \frac{1}{N}\sum_{i = 1}^{N}{\left(\frac{x_i - \bar{x}}{\sigma}\right)^4}$                                           
\\ \hline
\textit{Shape factor}   & $ X_{sf} = \frac{X_{rms}}{\frac{1}{N} \sum_{i=1}^{N}{|x_i|}}$  
\\ \hline
\textit{Crest factor}   & $ X_{cf} = \frac{max(|x_i|)}{X_{rms}} $                           \\ \hline
\textit{Impulse factor} & $ X_{if} = \frac{max(|x_i|)}{\frac{1}{N} \sum_{i=1}^{N}{|x_i|}} $                                                 
\\ \hline       
\textit{Clearance factor}  & $ X_{mf} = \frac{max(|x_i|)}{\left( \frac{1}{N} \sum_{i=1}^{N}{\sqrt{|x_i|}} \right)^2} $ 
\\ \hline
\textit{Average Amplitude Change (AAC)}        & $ X_{AAC} = \frac{1}{N}\sum_{i = 1}^{N - 1}{|x_{i + 1} - x_i|} $                             \\ \hline                                                                                     
\end{tabular}
\caption{Time-domain features}
\label{tab:td-features}
\end{table}

\subsection{Frequency-domain features}
The mechanical faults present themselves as oscillatory patterns which are combinations of frequencies with various amplitudes. The Fourier transform is one of the most prominent strategies in power spectral density estimation. Experts on vibrodiagnostics utilize it as a primary signal processing technique for data analysis as it is recommended in ISO 13373-2 standard~\cite{noauthor_iso_2016_2}.

@@

\begin{table}[ht]
\renewcommand{\arraystretch}{2}
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Feature}           & \textbf{Equation}                                                                                                  \\ \hline
\textit{Spectral centroid} & $ X_{fc} = \frac{\sum_{i = 0}^{N - 1}{f_i \cdot a(f_i)}}{\sum_{i = 0}^{N - 1}{a(f_i)}}$                   \\ \hline 
\textit{Energy}            & $ E(N) = \sum_{i = 1}^{N} a^2(t) $                                                                    \\ \hline \textit{Energy ratio}                & $E_r = E_i \;/\; \sum_{i = 1}^{N}{E_i} $                                                        \\ \hline
\textit{Spectral roll-on} & $ E(f_c) = 0.05 \cdot E(f_s / 2) $ \\ \hline                    
\textit{Spectral roll-off} & $ E(f_c) = 0.85 \cdot E(f_s / 2) $                                                   \\ \hline
\textit{Spectral flux}     & $X_{\mathrm{flux}} = 1 - \mathrm{corr}(A_{t-1}, A_t)$ \\ \hline 
\textit{Noisiness}                   & $X_{noise} = \mathrm{SNR}(X) = \mu \;/\; \sigma $                             \\ \hline
\textit{Shannon entropy}      & $H(X) = - \sum_{i} P(X = x_i) \cdot \ln(P(X = x_i)) $                                                  \\ \hline
\textit{Spectral negentropy}      & $\Delta I_E(f, \Delta f) = \sum_{k = 0}^{N - 1}{\frac{E(k, f, \Delta f)^2}{\frac{1}{N} \sum_{k=0}^{N-1} E(k, f, \Delta f)^2}} \cdot \ln\left(\frac{E(k, f, \Delta f)^2}{\frac{1}{N} \sum_{k=0}^{N-1} E(k, f, \Delta f)^2}\right)$                                                \\ \hline
\end{tabular}
\caption{Frequency-domain features}
\label{tab:fd-features}
\end{table}

In the frequency domain, we can obtain the usual statistical properties of the distribution being spectral centroid, skew and kurtosis. Additionally, spectral roll-on and roll-off, fundamental frequency, entropy, negentropy, spectral flux, signal-to-noise ratio (noisiness), energy in frequency bands, and energy ratio are extracted (Tab.~\ref{tab:fd-features})~\cite{peeters_large_2004}.

@@

\subsection{Bearing defect frequencies}
%TODO n = number of balls, d - average of outer ring diameter and outer ring, f_s - rotation speed, beta - contact angle - ball bearings close to 0

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{assets/analysis/bearing-parts.jpg}
    \caption{Bearing parts~\cite{lu_what_is_ball_bearing_2023}}
    \label{fig:bearing-parts}
\end{figure}

\begin{table}[h]
\centering
\renewcommand{\arraystretch}{2}
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Feature}            & \textbf{Equation}                                                                   
 \\ \hline
\textit{Outer race fault frequency} & $ BPFO = f_s \cdot \frac{n}{2} \cdot \left(1 - \frac{d}{D}\cos \beta \right)$               
\\ \hline
\textit{Inner race fault frequency} & $ BPFI = f_s \cdot \frac{n}{2} \cdot \left(1 + \frac{d}{D}\cos \beta \right)$
\\ \hline
\textit{Cage rotation frequency}    & $ FTF = f_s \cdot \frac{1}{2} \cdot \left(1 - \frac{d}{D}\cos \beta \right) $      
\\ \hline
\textit{Ball defect frequencies} & $ BSF = f_s \cdot \frac{D}{d} \cdot \left(1 + \left(\frac{D}{d}\cos \beta\right)^2 \right) $
\\ \hline
\end{tabular}
\caption{Bearing defect frequencies}
\label{tab:bearing-defect-features}
\end{table}

@@

\subsection{Feature importanance ranking}
Most of the scores are based on supervised learning, so they expect true class labels to apportion the measurements respectively. After the scores are assigned to the first $n$ features, those below a threshold are removed.

### 2024-04-27 thesis/chapters/analysis/3-diagnostics.tex
Export graphs from experiments and prepare outline of thesis

\textbf{Half-Space Tree} (HS-Tree) stands upon the concept of Isolation forest. It assumes that random splitting of each axis in the feature space will isolate outliers to their separate divisions sooner than non-deviant observations~\cite{gervasi_anomaly_2020,torres_automatic_2022}. This ensemble of trees is better suited for batch setting. HS-Tree stands out in adapting to changing streams because it is trained solely on normal data, requires constant memory, and is faster than density-based methods~\cite{tan_fast_2011}.

@@

The anomaly profile in the latest window is always compared to the predecessor reference window. After the latest window is filled up, it replaces the reference window. It suffices to use a window size of 250 and an ensemble of 25 trees~\cite{tan_fast_2011}.

>>>>

\textbf{Half-Space Tree} (HS-Tree) stands upon the concept of Isolation forest. It assumes that random splitting of each axis in the feature space will isolate outliers to their separate divisions sooner than non-deviant observations~\cite{gervasi_anomaly_2020, torres_automatic_2022}. This ensemble of trees is better suited for batch setting. HS-Tree stands out in adapting to changing streams because it is trained solely on normal data, requires constant memory, and is faster than density-based methods~\cite{tan_fast_2011}.

@@

### 2024-04-27 thesis/chapters/analysis/4-datasets.tex
Export graphs from experiments and prepare outline of thesis

\subsection{Machinery Fault Database}
MaFaulDa\footnote{\url{https://www02.smt.ufrj.br/~offshore/mfs/page_01.html}} is a collection of 1951 multivariate time series for 4 different operational conditions on rotor kit Alignment Balance Vibration Trainer (ABVT)~(Fig.~\ref{fig:mafaulda-simulator}). Each series has 5~seconds in duration and is captured at 50 kHz. Vibration signals were obtained with piezoelectric accelerometers with a linear response up to 10~kHz, amplitude range to $\pm$490 $\mathrm{m/s}^2$, and resolution step of 10.2~mV per $\mathrm{m/s}^2$.

@@

\subsection{CWRU bearings dataset}
In Case Western Reserve University (CWRU) bearing dataset\footnote{\url{https://engineering.case.edu/bearingdatacenter/download-data-file}} recordings were made of a fan end and drive end bearings under motor loads of 0, 1, 2, and 3~Horsepower (equivalently 0, 0.75, 1.49, 2.24~kW). Shaft speed was unaltered in all experiments, but it fluctuated between 1720 and 1797~rpm (approx. 29~Hz).

@@

\subsection{Unbalance of the rotating shaft}
Unbalance Detection of a Rotating Shaft\footnote{\url{https://www.kaggle.com/datasets/jishnukoliyadan/vibration-analysis-on-rotating-shaft}} is a Kaggle dataset that simulates 4 different unbalance strengths. 
\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{assets/analysis/rotating-shaft.jpg}
\caption{Motor driving shaft in unbalance measurement \cite{mey_machine_2020}}
\label{fig:rotating-shaft}
\end{figure}

>>>>

\subsection{Machinery Fault Database}
MaFaulDa~\cite{mafaulda_dataset} is a collection of 1951 multivariate time series for 4 different operational conditions on rotor kit Alignment Balance Vibration Trainer (ABVT)~(Fig.~\ref{fig:mafaulda-simulator}). Each series has 5~seconds in duration and is captured at 50 kHz. Vibration signals were obtained with piezoelectric accelerometers with a linear response up to 10~kHz, amplitude range to $\pm$490 $\mathrm{m/s}^2$, and resolution step of 10.2~mV per $\mathrm{m/s}^2$.

@@

\subsection{CWRU bearings dataset}
In Case Western Reserve University (CWRU) bearing dataset~\cite{cwru_dataset} recordings were made of a fan end and drive end bearings under motor loads of 0, 1, 2, and 3~Horsepower (equivalently 0, 0.75, 1.49, 2.24~kW). Shaft speed was unaltered in all experiments, but it fluctuated between 1720 and 1797~rpm (approx. 29~Hz).

@@

\subsection{Unbalance of the rotating shaft}
Unbalance Detection of a Rotating Shaft~\cite{unbalance_shaft} is a Kaggle dataset that simulates 4 different unbalance strengths. 
\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{assets/analysis/rotating-shaft.jpg}
\caption{Motor driving shaft in unbalance measurement \cite{mey_machine_2020}}
\label{fig:rotating-shaft}
\end{figure}
